### 2021-01-25 planet_paper/authors.tex

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

%
% key contributions
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\author[0000-0001-8732-6166]{J. D. Hartman}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

% PENDING
\author[0000-0001-7409-5688]{G. Stefansson} % halpha, bisectors, spec analysis
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%PENDING
\author[0000-0002-4891-3517]{G. Zhou} % metallicity analysis++
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

%PENDING
\author[0000-0002-9158-7315]{R. Brahm} %rafael.brahm@uai.cl coordination, WINE board
\affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

%PENDING
\author[0000-0002-5674-2404]{P. Evans}
\affiliation{El Sauce Observatory, Coquimbo Province, Chile}

%PENDING
\author[0000-0001-6588-9574]{K. A. Collins} % karenacollins@outlook.com
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

%PENDING
\author[0000-0001-8045-1765]{E. Flowers}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%PENDING
\author[0000-0002-8964-8377]{S. N. Quinn} % squinn@cfa.harvard.edu, CHIRON vetting spectrum
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}

% %PENDING
% \author{J. de Leon}
% \affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ky, Tokyo 113-0033, Japan}
% 
% %PENDING
% \author[0000-0002-4881-3620]{J. Livingston}
% \affiliation{Department of Astronomy, University of Tokyo, 7-3-1 Hongo, Bunkyo-ky, Tokyo 113-0033, Japan}
%
% %PENDING
% \author[0000-0003-2989-7774]{C. Bergmann}
% \affiliation{Exoplanetary Science at UNSW, School of Physics, UNSW Sydney, NSW 2052, Australia}
% \affiliation{Deutsches Zentrum f\"ur Luft- und Raumfahrt, M\"unchener Str. 20, 82234 Wessling, Germany}

%PENDING
\author[0000-0002-3481-9052]{K. G. Stassun}
\affiliation{Vanderbilt University, Department of Physics \& Astronomy, 6301 Stevenson Center Lane, Nashville, TN 37235, USA}
\affiliation{Fisk University, Department of Physics, 1000 17th Avenue N., Nashville, TN 37208, USA}
%

%
% Princeton team
%
%PENDING
\author[0000-0002-0628-0088]{W. Bhatti}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0001-7204-6727]{G. \'A. Bakos}
\affiliation{Department of Astrophysical Sciences, Princeton University, 4 Ivy Lane, Princeton, NJ 08540, USA}

%
% SG1 contributors
%

% % begin (ASTEP team)
% %PENDING
% \author{L. Abe} % Lyu
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %%PENDING->OMITTED
% %\author{A. Agabi} % Abdelkrim 
% %\affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author[0000-0001-7866-8738]{N. Crouzet} % Nicolas
% \affiliation{European Space Agency, European Space Research and Technology Centre (ESA/ESTEC), Keplerlaan 1, 2201 AZ Noordwijk, The Netherlands}
% 
% %PENDING
% \author[0000-0002-3937-630X]{G. Dransfield} % Georgina
% \affiliation{School of Physics \& Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
% 
% \author[0000-0002-7188-8428]{T. Guillot} % Tristan 
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author{W. Marie-Sainte} % Wenceslas
% \affiliation{Institut Paul \'{E}mile Victor, Concordia Station, Antarctica}
% 
% %PENDING
% \author{D. M\'ekarnia} % Djamel
% \affiliation{Universit\'e C\^{o}te d'Azur, Observatoire de la C\^ote d'Azur, CNRS, Laboratoire Lagrange, Bd de l'Observatoire, CS 34229, 06304 Nice cedex 4, France}
% 
% %PENDING
% \author[0000-0002-5510-8751]{A. H.M.J. Triaud} % Amaury
% \affiliation{School of Physics \& Astronomy, University of Birmingham, Edgbaston, Birmingham B15 2TT, United Kingdom}
% % end (ASTEP team)

%
% SG2/SG3 contributors
%
\author{J.~Teske} % PFS
\affiliation{Earth and Planets Laboratory, Carnegie Institution for Science, 5241 Broad Branch Road, NW, Washington, DC 20015, USA}

\author{S.~X.~Wang}% PFS
\affiliation{Department of Astronomy, Tsinghua University, Beijing 100084, People's Republic of China}

\author{R. P. Butler}% PFS
\affiliation{Earth and Planets Laboratory, Carnegie Institution for Science, 5241 Broad Branch Road, NW, Washington, DC 20015, USA}

\author{J. D. Crane}% PFS
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\author{S. A. Shectman}% PFS
\affiliation{Observatories of the Carnegie Institution for Science, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

% %PENDING
% \author[0000-0002-7595-0970]{C.~G.~Tinney}
% \affiliation{Exoplanetary Science at UNSW, School of Physics, UNSW Sydney, NSW 2052, Australia}

% %
% % WINE MPIA team
% %
% 
% %PENDING
% \author{T. Henning} %Thomas henning@mpia.de FEROS, WINE board
% \affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }
% 
% %PENDING
% \author[0000-0001-9513-1449]{N. Espinoza} %N\'estor nespinoza@stsci.edu WINE board
% \affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA}
% 
% %PENDING
% \author[0000-0002-5389-3944]{A. Jord\'an} %andres.jordan@gmail.com WINE board
% \affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
% \affiliation{Millennium Institute for Astrophysics, Chile}
% 
% %PENDING
% \author{M. Barbieri} %Maruo mauro.barbieri@uda.cl FEROS observations
% \affiliation{INCT, Universidad de Atacama, calle Copayapu 485, Copiap\'o, Atacama, Chile}
% 
% %PENDING
% \author{S. Nandakumar} %sangeetha.nandakumar@postgrados.uda.cl  FEROS observations
% \affiliation{INCT, Universidad de Atacama, calle Copayapu 485, Copiap\'o, Atacama, Chile}
% 
% %PENDING
% \author{T. Trifonov} %Trifon trifonov@mpia.de FEROS observations
% \affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }
% 
% %PENDING
% \author[0000-0002-1896-2377]{J.~I.~Vines} %Jose jose.vines.l@gmail.com FEROS observations
% \affiliation{Departamento de Astronom\'ia, Universidad de Chile, Camino El Observatorio 1515, Las Condes, Santiago, Chile}
% 
% %PENDING
% \author{M. Vuckovic} %Maja maja.vuckovic@uv.cl FEROS observations
% \affiliation{Instituto de F\'isica y Astronom\'ia, Universidad de Vapara\'iso, Casilla 5030, Valpara\'iso, Chile}

%
% SG4 contributors
%
%PENDING
\author[0000-0002-0619-7639]{C.~Ziegler} % carl.ziegler@dunlap.utoronto.ca
\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, Ontario M5S 3H4, Canada}

%%OMITTED
%\author{C.~Brice\~{n}o} % cbriceno@ctio.noao.edu
%\affiliation{Cerro Tololo Inter-American Observatory, Casilla 603, La Serena, Chile}

%PENDING
\author{N.~Law} % nlaw@unc.edu
\affiliation{Department of Physics and Astronomy, The University of North Carolina at Chapel Hill, Chapel Hill, NC 27599-3255, USA}

%PENDING
\author[0000-0003-3654-1602]{A.~W.~Mann} % awmann@unc.edu
\affiliation{Department of Physics and Astronomy, The University of North Carolina at Chapel Hill, Chapel Hill, NC 27599-3255, USA}

%PENDING
\author[0000-0002-2532-2853]{S. B. Howell} % non-detection, but write & inquire
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}

%PENDING
\author[0000-0001-7233-7508]{R. A. Matson}
\affiliation{U.S. Naval Observatory, Washington, DC 20392, USA}
%

% 
%-------------------------------------
% TESS Mission Architects:
% These authors should be listed in this order
% see https://spacebook.mit.edu/pages/viewpage.action?pageId=24543276
%-------------------------------------
%
%PENDING
\author{G. R. Ricker} % grr@space.mit.edu
\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0001-6763-6562]{R. Vanderspek} % roland@space.mit.edu
\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0001-9911-7388]{D. W.~Latham} % dlatham@cfa.harvard.edu
\affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
%
%PENDING
\author{S. Seager} % seager@mit.edu
\affiliation{Department of Earth, Atmospheric, and Planetary Sciences, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
%
%PENDING
\author[0000-0002-4715-9460]{J. M.~Jenkins} % jon.jenkins@nasa.gov
\affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}

%
%-------------------------------------
% 3 representatives of each of SPOC, POC, TSO, for a total of 9. 
%These 9 authors should be listed in alphabetical order
%-------------------------------------

% %PENDING
% \author[0000-0002-7754-9486]{C.~J.~Burke} % cjburke@mit.edu
% \affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %PENDING
% \author[0000-0003-2313-467X]{D.~Dragomir}
% \affiliation{Department of Physics and Astronomy, University of New Mexico, Albuquerque, NM, USA}
% 	
% %%PENDING->OMITTED
% %\author[0000-0003-0918-7484]{C.~X.~Huang}
% %\affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %%PENDING->OMITTED
% %\author{R.~C.~Kidwell, Jr.} % rkidwell@stsci.edu
% %\affiliation{Space Telescope Science Institute, 3700 San Martin Drive, Baltimore MD 21218 }
% 
% %PENDING
% \author[0000-0001-8172-0453]{A.~M.~Levine} % aml@space.mit.edu
% \affiliation{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% 
% %PENDING
% \author{E.~V.~Quintana} % elisa.quintana@nasa.gov
% \affiliation{NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
% 
% %PENDING
% \author[0000-0001-8812-0565]{J.~E.~Rodriguez}
% \affiliation{Center for Astrophysics \textbar \ Harvard \& Smithsonian, 60 Garden St, Cambridge, MA 02138, USA}
% 
% %PENDING
% \author[0000-0002-6148-7903]{J. C. Smith} % jeffrey.c.smith-1@nasa.gov
% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
% \affiliation{SETI Institute, Mountain View, CA 94043, USA}
% 
% %PENDING
% \author[0000-0002-5402-9613]{B. Wohler} % bill.wohler@nasa.gov
% \affiliation{NASA Ames Research Center, Moffett Field, CA 94035, USA}
% \affiliation{SETI Institute, Mountain View, CA 94043, USA}