### 2021-01-25 planet_paper/ms.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

\renewcommand*{\sectionautorefname}{Section} %for \autoref
\renewcommand*{\subsectionautorefname}{Section} %for \autoref

\newcommand{\tn}{TOI~1937} % target star name
\newcommand{\pn}{TOI~1937b} % planet name
\newcommand{\cn}{NGC~2516} % cluster name

\newcommand{\kms}{\,km\,s$^{-1}$}

\newcommand{\stscilink}{\textsc{\url{archive.stsci.edu/hlsp/cdips}}}
\newcommand{\datasetlink}{\textsc{\dataset[doi.org/10.17909/t9-ayd0-k727]{https://doi.org/10.17909/t9-ayd0-k727}}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{TBD.}%AAS journals.}
\shorttitle{TDB.}

\begin{document}

\defcitealias{bouma_wasp4b_2019}{B19}

\title{
  Cluster Difference Imaging Photometric Survey. III.
  Subtitle.
}

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

\begin{abstract}
  foo,bar.
\end{abstract}

\keywords{
	Exoplanets (498),
  Transits (1711),
	Exoplanet evolution (491),
	Stellar ages (1581),
	Young star clusters (1833)
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Lorem ipsum.

Section~\ref{sec:observations} describes the identification of the
candidate, and the follow-up observations that led to its confirmation.

The planet can only be understood with respect to its (putative)
host cluster, so in turn we analyze the
available six-dimensional positions and kinematics (Section~\ref{sec:gaia6d}),
the rotation periods of stars in \cn\ 
(Section~\ref{sec:rotation}), and the available lithium measurements
(Section~\ref{sec:lithium}).
We synthesize this data in
Section~\ref{sec:system} to present our best interpretation of the
system itself, in turn presenting our interpretation of the
cluster
(Section~\ref{subsec:cluster}), the star (Section~\ref{subsec:star})
and the planet (Section~\ref{subsec:planet}).  We conclude in
Section~\ref{sec:discussion} by discussing which questions we have
been able to answer, and which remain open.

\section{Identification and Follow-up Observations}
\label{sec:observations}

  \subsection{TESS Photometry}
  \label{subsec:tess}
  TESS: S7 + S9

  \subsection{Gaia Astrometry and Imaging}
  \label{subsec:gaia}
  Gaia:
    Within 20 arcsec, there are three sources.

  % 5489726768531119616 (target, G=13.02, plx=2.38 +/- 0.017 mas, so Bp-Rp=1.00).
  %   Note RV = 27.77km/s, +/- 9.99 km/s. Quotes E(Bp-Rp) = 0.1925

  % 5489726768531118848 (G=17.59, Bp=17.7, Rp=16.2, Bp-Rp=1.502,...
  %   but has plx=2.32 +/- 0.118 mas, and VERY SIMILAR proper motions. So, it's a
  %   binary. Presumably this is the companion that Ziegler found.)

  % There's also 5489726768531122560, a G=19.5 companion at further sep.

  % ...

  % Within 30 arcsec, there are... 8 sources. The others don't have parallaxes or
  % proper motions of significant interest.

  \subsection{High-Resolution Imaging}
  \label{subsec:speckle}
  Howell et al took some, for me, and had a non-detection.

  Also, Ziegler et al took some.
    They get the 2.1 arcsec separation, faint companion.

  \subsection{Ground-based Time-Series Photometric Follow-up}
  \label{subsec:groundphot}
  LCOGT
    5 light curves. 3 i band, 1 g band, 1 z band
  El Sauce
    1 r band.

  \subsection{Spectroscopic Follow-up}
  \label{subsec:spectra}

  \subsubsection{SMARTS 1.5$\,$m / CHIRON}
  \label{subsec:chiron}
  CHIRON x1 recon 2020-02-04, another circa 2021.

  \subsubsection{PFS}
  PFS template + RVs

  %\input{TOI837_rv_table.tex}

\section{Kinematics}

  \subsection{NGC 2516 broadly}
  \subsection{TOI 1937 specifically}

\section{Rotation}

  \subsection{NGC 2516 broadly}
  \subsection{TOI 1937 specifically}

\section{Lithium}

  \subsection{NGC 2516 broadly}
  \subsection{TOI 1937 specifically}

\section{System Modeling}
\label{sec:system}

\subsection{The Cluster}
\label{subsec:cluster}

\subsubsection{Physical Characteristics}
\label{subsec:clusterchar}

\subsubsection{HR Diagram}
\label{subsec:hr}

\subsection{The Star}
\label{subsec:star}

\subsubsection{Membership of \tn\ in \cn}
\label{subsec:member}

\subsubsection{Rotation}

\subsubsection{Lithium}

\subsubsection{Stellar Parameters}
\label{subsec:starparams}

\subsection{The Planet}
\label{subsec:planet}

\section{Discussion}
\label{sec:discussion}

The future:
- Detailed multi-abundance modelling.
  i.e,. Get spectra of other cluster stars using PFS, and compare the metallicities.
- 2$\farc$5 separate, multiband imager. (With LCOGT, 2m in Australia,
  z-band)->Do PSF fitting.
- K-band image. MagAO? Multicolor imager.
- Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\clearpage
\acknowledgements
\raggedbottom

The authors thank X and Y for fruitful discussions.
%
L.G.B. and J.H. acknowledge support by the TESS GI Program, program
NUMBER, through NASA grant NUMBER.
%
This study was based in part on observations at Cerro Tololo
Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop. ID
2020A-0146; 2020B-NUMBER PI: L{.}~Bouma), which is managed by the
Association of Universities for Research in Astronomy (AURA) under a
cooperative agreement with the National Science Foundation.
%
ACKNOWLEDGE PFS / CAMPANAS.
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%

% The ASTEP project acknowledges support from the French and Italian
% Polar Agencies, IPEV and PNRA, and from Universit\'e C\^ote d'Azur
% under Idex UCAJEDI (ANR-15-IDEX-01). We thank the dedicated staff at
% Concordia for their continuous presence and support throughout the
% Austral winter.
% %
% This research received funding from the European Research Council
% (ERC) under the European Union's Horizon 2020 research and innovation
% programme (grant n$^\circ$ 803193/BEBOP), and from the
% Science and Technology Facilities Council (STFC; grant n$^\circ$
% ST/S00193X/1).
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

This research was based in part on observations obtained at the
Southern Astrophysical Research (SOAR) telescope, which is a joint
project of the Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e
Inova\c{c}\~{o}es (MCTI/LNA) do Brasil, the US National Science
Foundation's NOIRLab, the University of North Carolina at Chapel Hill
(UNC), and Michigan State University (MSU).

This research made use of the Exoplanet Follow-up Observation
Program website, which is operated by the California Institute of
Technology, under contract with the National Aeronautics and Space
Administration under the Exoplanet Exploration Program.
% %
% This research made use of the NASA Exoplanet Archive, which is
% operated by the California Institute of Technology, under contract
% with the National Aeronautics and Space Administration under the
% Exoplanet Exploration Program.
% %

This research made use of the SVO Filter Profile Service
(\url{http://svo2.cab.inta-csic.es/theory/fps/}) supported from the Spanish
MINECO through grant AYA2017-84089.

Resources supporting this work were provided by the NASA High-End
Computing (HEC) Program through the NASA Advanced Supercomputing (NAS)
Division at Ames Research Center for the production of the SPOC data
products.
%

% A.J.\ and R.B.\ acknowledge support from project IC120009 ``Millennium
% Institute of Astrophysics (MAS)'' of the Millenium Science Initiative,
% Chilean Ministry of Economy. A.J.\ acknowledges additional support
% from FONDECYT project 1171208.  J.I.V\ acknowledges support from
% CONICYT-PFCHA/Doctorado Nacional-21191829.  R.B.\ acknowledges support
% from FONDECYT Post-doctoral Fellowship Project 3180246.
% %
% C.T.\ and C.B\ acknowledge support from Australian Research Council
% grants LE150100087, LE160100014, LE180100165, DP170103491 and
% DP190103688.
% %
% C.Z.\ is supported by a Dunlap Fellowship at the Dunlap Institute for
% Astronomy \& Astrophysics, funded through an endowment established by
% the Dunlap family and the University of Toronto.
% %
% D.D.\ acknowledges support through the TESS Guest Investigator Program
% Grant 80NSSC19K1727.
%
%
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  \texttt{arviz} \citep{arviz_2019},
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{ceres} \citep{brahm_2017_ceres},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  \texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001},
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
	%\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey,
    SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2}).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
	CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  PFS ({\bf CITE}),
  %  MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	AAT~(Veloce; \citealt{gilbert_veloce_2018}).
 	%Keck:I~(HIRES; \citealt{vogt_hires_1994}).
 	%{\bf VLT (number), UVES and GIRAFFE} (CITE: Pasquini et al 2002)
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

% \input{TOI837_phot_table.tex}
% \input{TOI837_rv_table.tex}
% \input{ic2602_ages.tex}
\input{starparams.tex}
\input{compparams.tex}
% \input{model_posterior_table.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\listofchanges

%\allauthors
\end{document}